Hybrid recurrent model for regime switching nonstationary time series

### 2021-10-01 iclr2022/hybridmodel_iclr.tex
add results on 1000ts

\newcommand{\ts}{y}
\newcommand{\fullts}{{\bf \ts}}
\newcommand{\tspred}{\hat{\ts}}
\newcommand{\stat}{f}
\newcommand{\statparam}{\theta_{stat}}
\newcommand{\fullstat}{{\bf \stat}}
\newcommand{\lag}{h}
\newcommand{\window}{w}
\newcommand{\tswindow}{{\bf \ts}}
\newcommand{\meants}{\Bar{\ts}}
\newcommand{\rnnwindow}{{\bf \rnn}}
\newcommand{\rnn}{z}
\newcommand{\err}{err}
\newcommand{\errwindow}{{\bf \err}}
\newcommand{\rnnmodel}{\textsc{rnn}}
\newcommand{\ws}{w}
\newcommand{\fullws}{{\bf \ws}}
\newcommand{\wswindow}{{\bf \ws}}
\newcommand{\concatinput}{x}
\newcommand{\fullconcatinput}{{ \bf \concatinput}}
\newcommand{\numberts}{10000}
\newcommand{\threshold}{\eta}
\newcommand{\predictor}{\mathrm{RNN}_p}
\newcommand{\classifier}{\mathrm{RNN}_c}
\newcommand{\rnnparam}{\theta}
\newcommand{\remainder}{r}
\newcommand{\hiddenregime}{U}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \citep{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
 Parametric generative models allow to provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \citep{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685} or the ARIMA with the Box-Jenkins approach \citep{box2015time} are for instance very popular parametric linear generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data, possibly lying in a high dimensional state space, describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain (or a mixing process)  the observations arising therefrom present a highly complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
 In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \citep{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \citep{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Altough these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

In this paper, we consider  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ sequences representing the apparition of fashion trends on social media per week from 01-01-2015 to 01-01-2019. Our fashion time series has very appealing properties: they have all the same length, same seasonality, no missing value and the absence of sparse time series even for niche trends. The originality of our dataset come from the fact that additional external weak signals can be introduced. With our fashion expertise, we detected several groups of fashion users with a high influence power. Analysing their specific behaviours on social medias, we add to each time series, 4 external weak signals representing the same fashion trends but on a sub-category of users. We call them weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with our large fashion dataset, leverage our complex external weak signals and finally provide the most accurate forecasts.

Recurrent neural network are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining DNN architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
%However, the hybrid model has not been designed to deal with external signals even if it include a neural network.

@@

\section{Hybrid model with external signals}
\label{sec:hybrid}
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global RNN. We use a per time series statistical model to learn local behaviour and to normalize sequences before feeding them to the neural network.
We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global neural network framework composed of two RNN. We use a per time series statistical model to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, we use two recurrent neural networks and weak signals to correct, if needed, the predictions provided by the first collection of parametric models.

@@

\subsection{Per time series statistical model}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
%The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series).
The time-series-specific models compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n_{\statparam^n}$ the $n$-th parametric model of the the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{stat,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{stat,n}_{T+i|T} = \stat^n_{\statparam^n}(\ts^n_{1:T})_i\,.
$$

During the M4 competition, the hybrid model of \cite{} is based on a multiplicative exponential smoothing model as the time-series-specific model. However, on sporadic time series, this choice could lead to poor result and instability. In our approach, we decided to go deeper and provide a general framework able to deal with any kind of per-time-series models. In Section~\ref{sec:exp}, we present two versions of our framework. The first one use an  exponential smoothing as reference  similar to \cite{} and the second one use a TBATS model \cite{} which provides better results. The main complexity is in the next stage: training a global RNN to correct the per-time-series statistical models when the statistical models provide poor performance.
% Indeed, depending on time series, some family of parametric models can achieve really heterogeneous level of accuracy. 
%For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n, \lvl^n_0, \{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters, $\lvl^n_0$ the first level term and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. Given the sequence $\fullts^n$ and the parameters $\Theta^n$ , the two components  $\fulllvl^n$ and $\fullseas^n$ are computed as follow :
%We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
%\begin{align*}
    %\lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    %\seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
%\end{align*}

@@

%At this stage, we want to train a global RNN, trained on all the N sequences and able to understand and correct the weaknesses of the first per-time-series parametric models.
%At this stage, we can write that for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
%$$
%\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \varepsilon^n_{T+i}\,,
%$$
In this paper, we introduce a model based on a latent regime. In regime 0, the first parametric forecast succeeds at catching all the information of the $n$-th sequence and the final forecast of our framework is the prediction of the per-time-series parametric model. Conversely, in regime 1, the first parametric forecast is unaccurate and it remains a residual random error term that can be estimated. % quantity that we note $\remainder$. 
Conditionally on $\{\fullts^n\}_{1\leqslant n\leqslant N}$, our regime-switching model is:

$$
\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \hiddenregime^n_T \cdot \remainder^{n}_{T+i} + \varepsilon^n_{T+i} = \stat^n_{\statparam^n}(\ts^n_{1:T})_i + \hiddenregime^n_T \cdot \remainder^{n}_{T+i} + \varepsilon^n_{T+i}\,,
$$
where $\{\varepsilon^n_{T+i}\}_{1\leqslant i\leqslant \lag}$ are independent random variables, $U^n_T$ is an unobserved regime switching Bernoulli random variable and $\{\remainder^n_{T+i}\}_{1\leqslant i\leqslant \lag}$ is the random error term to be estimated.
%In the regime 1, the first parametric forecast is unaccurate and it remains a non-forecasted quantity that we note $\remainder$.
%For this regime, we have for all $i \in \{1,\ldots,\lag\}$ :
%$$
%\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \remainder^{n}_{T+i} + \varepsilon^n_{T+i}\,,
%$$

To estimate this model, we introduce our new regime swithing error-corrector framework. We train two global RNN on all sequences, a classifier ($\classifier$) and a predictor ($\predictor$). We train the classifier to estimate the law of the Bernoulli random variable $\hiddenregime^n_T$ we train the predictor to predict $\{\remainder^{n}_{T+i}\}_{1\leqslant i\leqslant \lag}$. However, these two tasks require a thorough data pre-processing and with our collection of parametric models, we propose first to normalize and deseasonalize the RNNs inputs.

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is equal to the forecast horizon $\window = \lag$. Consider the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnn^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnn^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{stat,n}_{T-w+i|T-w}}{\meants^n_T}\,,
$$
where $ \meants^n_T = \sum_{i = 1}^{w}  \ts^n_{T-w+i}/w$.
We use this sequence as input of  $\classifier$ and $\predictor$. %Write, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,H\}$, $\err^{n,T}_{T+i|T} = \ts^n_{T+i} - \etspred^{ets,n}_{T+i|T}$.
 Given $\rnnwindow^n_T$, the proposed regime swithing error-corrector reccurent model is given, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$, by :
\begin{align*}
\ts^n_{T+i}  &= \tspred^{stat,n}_{T-w+i|T-w} +  \hat{\hiddenregime}^n_T * \hat{\remainder}^{n}_{T+i} + \varepsilon^n_{T+i}\,, \\
\hat{\hiddenregime}^n_T &= \mathbf{1}_{\classifier(\rnnwindow^n_T) > \threshold}\,, \\
\hat{\remainder}^{n}_{T+i} &= \predictor(\rnnwindow^n_T)_{T+i} \times \meants^n_T\,.
\end{align*}

where $\mathbf{1}$ is the indicator function and $\threshold$ a threshold parameter.

@@

Using a time-series-specific statistical model, the hybrid network achieves striking accuracy on single sequences. Then, the shared recurrent neural network relies on the strength of machine learning methods at leveraging a huge quantities of heterogeneous data.
In addition to the $N$ target time series, we consider $K \times N$ external sequences indexed from $0$ to $T$. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ linked to the sequence $\fullts^n$. We note $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ all the values of the family of weak signals linked to the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the family. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, we define a new input for our two RNN,   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnn^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ our final model becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{equation}
\label{eq:full:model}
\ts^n_{T+i}  = \tspred^{stat,n}_{T-w+i|T-w} +  \mathbf{1}_{\classifier(\fullconcatinput^n_T) > \threshold} \times \predictor(\fullconcatinput^n_T)_{T+i} \times \meants^n_T + \varepsilon^n_{T+i}\,,
\end{equation}
An illustration of our proposed architecture is displayed in Figure~\ref{fig:architecture}.

\begin{figure}
  \centering
    \includegraphics[width=1\linewidth]{figure/Regime_switching_recurrent_model}
  \caption{Hybrid model with weak signals architecture.}
\label{fig:architecture}
\end{figure}

@@

As the regime is not observed, the loglikelihood of the observations cannot be written explicitly to train the model. Therefore we use Fisher's identity to compute the score function and write a gradient ascent algorithm to estimate simultaneously the weights of RNN-p and of RNN-c denoted by $\rnnparam = \{\rnnparam_c,\rnnparam_p\}$. The score is
\begin{equation}
\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h}|\ts^n_{1:T}) = \mathbb{E}\left[\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T})\middle|\ts^n_{1:T+h}\right]\,,
\end{equation}
\\
where the joint likelihood $p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T}) = p_\theta(U_T^n|\ts^n_{1:T})p_\theta(\ts^n_{T+1:T+h}|U_T^n,\ts^n_{1:T})$ is given by \eqref{eq:full:model} and the expectation is under the posterior distribution of $U_T^n$ given $\ts^n_{1:T+h}$.
A complete description of the training process can be found in Appendix.

>>>>

\newcommand{\ts}{y}
\newcommand{\fullts}{{\bf \ts}}
\newcommand{\tspred}{\hat{\ts}}
\newcommand{\stat}{f}
\newcommand{\statparam}{\theta_{stat}}
\newcommand{\fullstat}{{\bf \stat}}
\newcommand{\lag}{h}
\newcommand{\window}{w}
\newcommand{\tswindow}{{\bf \ts}}
\newcommand{\meants}{\Bar{\ts}}
\newcommand{\rnnwindow}{{\bf \rnn}}
\newcommand{\rnn}{z}
\newcommand{\err}{err}
\newcommand{\errwindow}{{\bf \err}}
\newcommand{\rnnmodel}{\textsc{rnn}}
\newcommand{\ws}{w}
\newcommand{\fullws}{{\bf \ws}}
\newcommand{\wswindow}{{\bf \ws}}
\newcommand{\concatinput}{x}
\newcommand{\fullconcatinput}{{ \bf \concatinput}}
\newcommand{\numberts}{10000}
\newcommand{\threshold}{\eta}
\newcommand{\predictor}{RNN_p}
\newcommand{\classifier}{RNN_c}
\newcommand{\rnnparam}{\theta}
\newcommand{\remainder}{r}
\newcommand{\hiddenregime}{U}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \citep{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
 Parametric generative models allow to provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \citep{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685} or the ARIMA with the Box-Jenkins approach \citep{box2015time} are for instance very popular parametric linear generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data, possibly lying in a high dimensional state space, describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain (or a mixing process)  the observations arising therefrom present a highly complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
 In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \citep{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \citep{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Altough these works provide promising results, HMM are computationally expensive to train and are not yet well studied for sequences with thousands of components with a seasonal behavior \textcolor{blue}{with thousands of components or with a seasonal behavior ?}.

@@

In this paper, we consider  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ sequences representing the apparition of fashion trends on social media per week from 01-01-2015 to 01-01-2019. Our fashion time series has very appealing properties: they have all the same length, same seasonality, no missing value and the absence of sparse time series even for niche trends. The originality of our dataset come from the fact that additional external weak signals can be introduced. With our fashion expertise, we detected several groups of fashion users with a high influence power. Analysing their specific behaviours on social medias, we add to each time series, 4 external weak signals representing the same fashion trends but on a sub-category of users. We call them weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at showing a model able to deal with our large fashion dataset, leverage our complex external weak signals and finally provide the most accurate forecasts.

Recurrent neural network seems to be the perfect candidate to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining DNN architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
However, the hybrid model has not been designed to deal with external signals even if it include a neural network.

@@

\section{Hybrid model with external signals}
\label{sec:hybrid}
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global RNN. We use a per time series statistical model to learn local behaviour and to normalize sequences before feeding them to the neural network.
We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global neural network framework composed of 2 RNN. We use a per time series statistical model to learn local behaviour, to normalize sequences and to compute a first forecast. Then, we use 2 recurrent neural networks and weak signals to correct, if needed, the predictions provided by the first collection of parametric models.

@@

\subsection{Per time series statistical model}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
%The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series).
The time-series-specific models compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n$ the $n$-th parametric model linked to the $n$-th sequence and $\statparam^n$ its parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$, the parametric models $\{\stat^n\}_{1\leqslant n \leqslant N}$ and its associated parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{stat,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{stat,n}_{T+i|T} = \stat^n(\ts^n_{1:T}|\statparam^n)_i\,,
$$

During the M4 competition, the hybrid model's author based his framework on a multiplicative exponential smoothing model as the time-series-specific model. However, on sporadic time series, this choice could lead to poor result and instability. In our approach, we decided to go deeper and provide a general framework able to deal with any kind of per-time-series models. In the Section~\ref{sec:exp}, we present two versions of our framework. The first one use an additive exponential smoothing as reference and the second one use a TBATS model as reference. At the end, The main complexity is in the next stage: succeed in training a global RNN to understand all the per-time-series statistical models and correct their mistakes if needed.
% Indeed, depending on time series, some family of parametric models can achieve really heterogeneous level of accuracy. 
%For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n, \lvl^n_0, \{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters, $\lvl^n_0$ the first level term and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. Given the sequence $\fullts^n$ and the parameters $\Theta^n$ , the two components  $\fulllvl^n$ and $\fullseas^n$ are computed as follow :
%We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
%\begin{align*}
    %\lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    %\seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
%\end{align*}

@@

%At this stage, we want to train a global RNN, trained on all the N sequences and able to understand and correct the weaknesses of the first per-time-series parametric models.
%At this stage, we can write that for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
%$$
%\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \varepsilon^n_{T+i}\,,
%$$
At this stage, we assume that we have two different regimes. In the regime 0, the first parametric forecast succeeds at catching all the information of the $n$-th sequence. the final forecast of our framework is the prediction of the per-time-series parametric model. Conversely, in the regime 1, the first parametric forecast is unaccurate and it remains a non-forecasted quantity that we note $\remainder$. Our regime-switching system is then :

$$
\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \hiddenregime^n_T * \remainder^{n}_{T+i} + \varepsilon^n_{T+i}\,,
$$
where $\{\varepsilon^n_{T+i}\}_{i\leqslant 1}$ are independent random variables, $\{\remainder^n_{T+i}\}_{i\leqslant 1}$ are non null remainders and $U^n_T$ an unobserved regime switching Bernoulli random variable.
%In the regime 1, the first parametric forecast is unaccurate and it remains a non-forecasted quantity that we note $\remainder$.
%For this regime, we have for all $i \in \{1,\ldots,\lag\}$ :
%$$
%\ts^n_{T+i} = \tspred^{stat,n}_{T+i|T} + \remainder^{n}_{T+i} + \varepsilon^n_{T+i}\,,
%$$

To estimate this system, we introduce our new regime swithing error-corrector framework. We train two global RNN on all sequences, a classifier ($\classifier$) and a predictor ($\predictor$). We train the classifier to provide a probability of being in the regime 1 and we train the predictor to estimate the first per-time-series forecasts errors if we are in the regime 1. However, these 2 tasks require a thorough data pre-processing and with our collection of parametric models, we propose a fine way to normalize and deseasonalize the RNNs inputs.

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is equal to the forecast horizon $\window = \lag$. Consider the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnn^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnn^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{stat,n}_{T-w+i|T-w}}{\meants^n_T}\,,
$$
where $ \meants^n_T = \sum_{i = 1}^{w}  \ts^n_{T-w+i}/w$.
We use this sequence as input of our two neural network, $\classifier$ and $\predictor$. %Write, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,H\}$, $\err^{n,T}_{T+i|T} = \ts^n_{T+i} - \etspred^{ets,n}_{T+i|T}$.

Given $\rnnwindow^n_T$, the proposed regime swithing error-corrector reccurent model is given, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$, by :
\begin{align*}
\ts^n_{T+i}  &= \tspred^{stat,n}_{T-w+i|T-w} +  \hat{\hiddenregime}^n_T * \hat{\remainder}^{n}_{T+i} + \varepsilon^n_{T+i} \\
\hat{\hiddenregime}^n_T &= \mathbf{1}_{\classifier(\rnnwindow^n_T) > \threshold} \\
\hat{\remainder}^{n}_{T+i} &= \predictor(\rnnwindow^n_T)_{T+i} \times \meants^n_T
\end{align*}

where $\mathbf{1}$ is the indicatrice function and $\threshold$ a threshold parameter.\\

@@

Using a time-series-specific statistical model, the hybrid network achieves striking accuracy on single sequences. Then, the shared recurrent neural network relies on the strength of machine learning methods at leveraging a huge quantities of heterogeneous data.
In addition to the $N$ target time series, we consider $K \times N$ external sequences indexed from $0$ to $T$. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ linked to the sequence $\fullts^n$. We note $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ all the values of the family of weak signals linked to the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the family. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, we define a new input for our two RNN,   
\begin{align*}
\fullconcatinput^n_T &= \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}\, \\
&= \{\rnn^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,
\end{align*}

Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ our final model becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{equation}
\label{eq:full:model}
\ts^n_{T+i}  = \tspred^{stat,n}_{T-w+i|T-w} +  \mathbf{1}_{\classifier(\fullconcatinput^n_T) > \threshold} \times \predictor(\fullconcatinput^n_T)_{T+i} \times \meants^n_T + \varepsilon^n_{T+i}\,,
\end{equation}
An illustration of our proposed architecture is displayed in Figure~\ref{fig:architecture}.

\begin{figure}
  \centering
    \includegraphics[width=1.2\linewidth]{figure/Regime_switching_recurrent_model}
  \caption{Hybrid model with weak signals architecture.}
\label{fig:architecture}
\end{figure}

@@

As $\hiddenregime$ is not observed, the loglikelihood of the observations cannot be written explicitly to train the model. Therefore we use Fisher's identity to compute the score function and write a gradient ascent algorithm to estimate simultaneously the weights of RNN-p and of RNN-c denoted by $\rnnparam = \{\rnnparam_c,\rnnparam_p\}$. The score is
\begin{equation}
\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h}|\ts^n_{1:T}) = \mathbb{E}\left[\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T})\middle|\ts^n_{1:T+h}\right]\,,
\end{equation}
\\
where the joint likelihood $p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T}) = p_\theta(U_T^n|\ts^n_{1:T})p_\theta(\ts^n_{T+1:T+h}|U_T^n,\ts^n_{1:T})$ is given by \eqref{eq:full:model} and the expectation is under the posterior distribution of $U_T^n$ given $\ts^n_{1:T+h}$.
A complete description of the training process can be found in Appendix.
